Novel genetic control of migratory diapause in Australian Monarch Butterflies

### 2023-04-10 paper/paper_draft.Rmd
paper updates with some of micah's comments

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(snpR); library(data.table); library(ggplot2); library(gridExtra); library(dplyr); library(openxlsx); library(lme4)

@@

Migratory species are acutely at risk from global anthropogenic environmental change, since they rely on multiple distinct ecosystems and consistent seasonal timings [@Both2010]. As a result, the loss of breeding, wintering, or any transitional habitats may all cause declines in migrants [@Robbins1989], as may shifts in resource phenology due to seasonal changes which can cause resource availability and species life-histories to fall out of synch [@Jones2010a]. It is therefore not surprising that widespread declines have been observed in many migratory taxa [@Limburg2009;@Robbins1989].

Population-level adaptability in migratory life-history should therefore be expected to increase long-term population viability by conferring resistance to phenological and environmental changes [@Pulido7341]. Migratory life-history adaptation can essentially occur along three distinct axis: 1) *temporal*, or adaptation that changes the phenology of migration; 2) *spatial*, or adaptation that alters migratory pathways and orientation; and 3) *residency*, or adaptation that shifts populations between migratory and resident life histories. Inter-population variation and adaptation in each of these pathways is well known from nature: temporal changes or variation in migration are well documented in many taxa, particularly in birds [@Rubolini2007; @Gordo2007]; short-stopping, extension of migratory pathways, or complete re-direction of migration is well known in birds [@Ambrosini2011; @LaSorte2007; @Visser2009]; and shifts towards residency or partial migration have been observed in birds [@Pulido7341; @Plummer2015; @Møller2014], butterflies [@Howard2010], mammals [@VanDerRee2006; @Fleming2003; @Jones2014a], and other taxa.

Adaptation via a transition to residency is of both particular interest and concern for the conservation of migratory species. While switching to a resident life history can allow species to avoid the increasing risks of migration [and the intrinsically very energetic large costs of migration in general, see @McWilliams2004; @Rankin1992; @Wikelski2003], there is reasonable concern that the "extinction" of migration may not be easily reversible in some species. This is concerning since migratory life-history variation is expected to increase population adaptability [@Pulido2007], and thus the permanent loss of migration could potentially compromise the long term population persistence.

This is, for example, a major concern for Monarch Butterflies (Danaus plexippus). North American migratory monarch populations have experienced severe declines over the last several decades, sparking a fear of what Semmens et al. [-@Semmens2016] called a "Quasi-extinction" of the migratory phenotype driven by changes in the selective forces acting on the species to favor resident individuals. In addition to the damage to the long-term population persistence that could result from the loss of migration in North American, the annual monarch migration is publicly-beloved and its loss would be culturally costly [@Gustafsson2015].

Interestingly, however, there is at least one population of Monarchs in Australia in which migratory behavior has re-emerged after previously shifting to a resident life history. This population, which was established sometime before 1871, is the product of a series of sequential self-introductions that stepping-stoned across the Pacific ocean from North America beginning sometime in the early 1840s after the introduction of their host plants enabled individuals to survive after being blown from the mainland [@Zalucki2004]. Monarchs are not migratory on any Pacific islands, and no migratory behavior was observed until the 1930s, when non-reproductive overwintering aggregations were observed in northern New South Wales [NSW, see @James2019]. These and other individuals were later shown to migrate hundreds of kilometers each year to their wintering grounds from more southerly, inland locations in NSW [@James2019]. After over 50 years (or hundreds of generations) without migration, a migratory life-history re-emerged in these populations. Understanding how this happened could provide crucial context around the "extinction" of migration in monarchs and other species.

To better understand how migration re-emerged in Australian monarchs, we studied continuously breeding monarch populations in Queensland (QLD). While these populations do not experience substantial seasonal fluctuations in resources availability, and thus do not migrate [@James2019], recent work has shown that reproductive diapause can be induced in some, *but not all* individuals from these populations if they are exposed to a reduction in photoperiod during development [@Freedman2018]. In brief, reproductive diapause, where individuals delay their reproductive maturation, is a critical phenotype for migratory individuals because it allows them to invest more heavily in energy storage and increase their average longevity [@Herman2001]. Like other migratory behaviors, diapause is controlled by changes in day length, temperature, and other seasonal fluctuations [@Goehring2002].

The variance in diapause induction in QLD monarchs allows for a direct search for genetic associations with the trait and thus for an examination of the genetic causes of their re-evolution of migration. *Here, we used a mix of reduced representation and whole genome sequencing to search for genes associated with migratory diapause in QLD monarch butterflies. We found one strongly associated locus which had previously been identified to be associated with juvenile phase transitions in silk moths (insert species name), which is controlled by the same major hormone as reproductive diapause. This gene had not previously been found to be associated with migration in North American monarch butterflies, suggesting that diapause and potentially migration as a whole was secondarily re-evolved in Australian monarchs by novel mutations.*

# Methods
## Sample Collection, Incubation, and Phenotyping
```{r, read_sample_meta_get_numbers, echo=FALSE,message=FALSE,include=FALSE,warning=FALSE}
sample.meta <- read.table("../data/sample_metadata.txt", header = T, stringsAsFactors = F)
old.data <- data.table::fread("../data/Aus_2018_metadata.txt", header = T)
old.data[,fam := as.numeric(gsub("_.+", "", ID))]

@@

We collected 22 female monarch butterflies from Pinjarra Hills, Queensland, Australia (27°32′26.7″S, 152°54′22.7″E) between the 5th and 9th of July 2018. We noted individuals in all life stages present at the field site during our sample collections. We then enclosed these individuals on their host plants and collected the eggs produced by 21 of them. These eggs represent 21 total material families, but, since females may mate multiple times and lay clutches with mixed parentage [CITE], each of these families may consist of multiple groups of half-siblings. We did not remove *Ophryocystis elektroscirrha* (a protozoan parasite) spores, which were uniformly present on our eggs [consistent with previous reports from this location, see @Freedman2018].

<!-- add citation -->

We then incubated all of the eggs using the same incubators and following the "decreasing photoperiod" experimental scheme of Freedman et al. [-@Freedman2018] in order to induce reproductive diapause and ensure that our data could be later pooled with that produced in the earlier study. We released all male butterflies following emergence, since phenotyping diapause in male monarchs can be challenging due to the difficulty in extracting and weighing the seminal vesicle, which is likely the best indicator of male reproductive development [@Freedman2018]. This left us with a total of `r nrow(old.data)` female individuals from `r length(unique(na.omit(old.data$fam)))` maternal families.

We then assessed reproductive diapause by assessing the degree of oocyte development in each of our adult female butterflies. We observed that individuals fell into two clearly defined bins: `r sum(old.data$Phenotype == "C", na.rm = TRUE)` out of `r nrow(old.data)` had fully developed, chorionated eggs (visible ridges along the exterior of the egg), `r sum(old.data$Phenotype != "C", na.rm = TRUE)` females did not. Of these latter individuals a few (`r sum(old.data$Phenotype == "I", na.rm = TRUE) + sum(old.data$Phenotype == "N", na.rm = TRUE)`) had no yolk in their eggs whatsoever, but most had some degree of yolking. Since vitellogenesis generally only begins to a substantial degree after eclosion in monarchs [@Pan1976], We called any individuals with more than a small amount of yolking in their eggs as reproductively mature (N = `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "M", na.rm = TRUE)` and `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "R", na.rm = TRUE)` for reproductively immature and mature individuals, respectively, see Supplementary Information for images of egg development in each individual). Wings and bodies from all individuals were preserved in dry coin envelops or 95% ethanol, respectively.

We supplemented these samples with preserved, dry butterfly wings from each of the "decreasing photoperiod" female individuals previously phenotyped for reproductive development by Freedman et al. [-@Freedman2018], constituting an additional `r sum(sample.meta$year == 2016)` total female butterflies from `r length(unique(meta.micah$Female_Line[meta.micah$Treatment == "Decreasing"]))` maternal families, of which `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "M", na.rm = TRUE)` and `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "R", na.rm = TRUE)` were phenotyped by the authors as reproductively immature and mature, respectively. Together, these two datasets contained `r length(na.omit(sample.meta$likely_pheno))` monarch butterflies, of which `r sum(sample.meta$likely_pheno == "M", na.rm = TRUE)` and `r sum(sample.meta$likely_pheno == "R", na.rm = TRUE)` were reproductively immature and mature, respectively (see Table 1).

## Sequencing and Genotyping
We removed and extracted DNA one leg from each of our samples collected in 2018 and a wing base from each of the samples collected by Freedman et al [-@Freedman2018] using the magnetic bead protocol of Ali et al [-@Ali389]. We quantified the resulting DNA on a BioTek Instruments FLx800 Fluorescence Reader using Thermo Fisher Scientific Quant-iT PicoGreen dsDNA Reagent, the prepared Restriction Associated Digest (RAD) libraries using the Pst1 restriction enzyme according to Ali et al [-@Ali389]. We sequenced these 150 bp paired-end sequencing libraries using an Illumina Hi-Seq 4000.

We aligned the resulting raw sequence data to the v4 monarch butterfly genome assembly [@Gu2019] using the mem algorithm of the Burrows-Wheeler Aligner [@Li2009a]. We then filtered out PCR duplicates, improperly paired, or poorly mapped reads using SAMtools [@Li2009].

@@

<!-- and then used the ngsParalog tool to removed an potentially paralogous loci by restricting our analysis to loci more than a kilobase away from loci with a log ratio test statistic > 10 [@Linderoth2018]. -->

<!-- Didn't do paralog, redoing now did do a rough HWE filter. Probably shouldn't really do paralog check, though, since there is a lot of family structure in our samples, which will throw things off. -->

We then called gentoypes using the ANGSD software package with the following parameters: -doMajorMinor 1 (determine major and minor alleles using a genotype likelihood approach),  -doMaf 2 (determine minor allele frequencies), -SNP_pval 1e-8 (keep only loci with a SNP p-value $\le 1 \times 10^{-8}$), -doGeno 4 (call genotypes), -doPost 2 (calculate genotype posterior probabilities using a uniform prior), -postCutoff 0.95 (keep only loci where the highest genotype posterior probability $ \ge 0.95$), -minQ 20 (keep only loci with a sequencing quality $\ge 20$), -minMapQ 20 (keep only loci with a mapping quality $\ge 20$), -minInd 130 (keep only loci sequenced in at least 130 individuals), and -minMaf 0.05 (keep only loci with a minor allele frequency $\ge 0.05$).

@@

```{r import_and_diversity, echo=FALSE, message=FALSE, warning=FALSE,include=FALSE}
# L <- # total number of sequenced bases

@@

#==========diversity===========
# dat <- calc_pi(dat, c(".base", "year"))
# dat <- calc_ho(dat, c(".base", "year"))
# dat <- calc_fis(dat, c(".base", "year"))
# dat <- calc_pairwise_fst(dat, "year", boot = 100, boot_par = 6, verbose = TRUE)
# saveRDS(dat, "diversity_stats.RDS")
dat <- readRDS("diversity_stats.RDS")
diversity <- get.snpR.stats(dat, c(".base", "year"), c("pi", "ho", "fis"))$weighted.means
fst <- get.snpR.stats(dat, "year", "fst")$weighted.means

@@

In order to determine if any genomic regions were significantly associated with We used the snpR R package to filter out loci significantly out of Hardy-Weinburg Equilibrium ($p \le 0.000001$) in either year in order to help remove any paralogous loci and interpolated missing data by conducting binomial draws at missing genotypes using the minor allele frequencies at those loci [@Hemstrom2021]. To assess the basic level of genetic diversity amongst our samples, we then calculated the average number of pairwise differences ($\pi$), observed heterozygosity ($H_{O}$), and $F_{IS}$ according to Weir and Cockerham [-@Weir1984] for all of our samples and for each year independently using snpR [@Hemstrom2021]. We also calculated $F_{ST}$ according to Weir and Cockerham [-@Weir1984] between sample years and between both sample year and reproductive phenotype jointly[@Hemstrom2021]. We bootstrapped a significance level for $F_{ST}$ by bootstrapping individuals among populations with replacement 100 times. To determine if there was any substantial unexpected genetic structuring (other than family groups) in our samples, we additionally both conducted a Principal Component Analysis using snpR [@Hemstrom2021]. Since some individuals escaped their rearing dishes (but not the incubator) during rearing, we also confirmed the parentage of our samples using the Colony2 program, assuming polygamous, random mating [@Jones2010].

@@

# plot
Figure2b <- plot_manhattan(dat, "gmmat_pval_likely_pheno", chr = "scaffold", significant = 0.00005, suggestive = 0.0005, log.p = T,
                           chr_order = c(paste0("chr", 1:30), "other"), highlight_style = "red") # highlight top snp
Figure2b <- Figure2b$plot + ylab("-log10(p-value)") + xlab("Position") + theme(strip.text = element_blank())
```
In order to determine the degree to which diapause status was biased between families, we constructed a Fisher's exact test for independence using 2000 simulations in R. We then conducted a Genome-Wide Association Study (GWAS) in order to determine if any loci were significantly associated with reproductive status. Since our samples were composed of many groups of full and half-siblings and came from two genetically different years, we used the GMMAT R package to conduct a population and family structure-corrected GWAS [@Chen2020; @Chen2016]. We visualized the results by constructing a Manhattan and quantile-quantile plot using snpR [@Hemstrom2021]. We identified the genes underlying any candidate regions by hand using the MonarchBase genome browser online tool [@Zhan2013].

@@

We also briefly re-analyzed phenotypic data from the same earlier study that provided our 2016 monarch studies [@Freedman2018] in order to determine if wing shape or size was correlated with reproductive diapause by fitting a linear mixed effect model with wing shape or size as response variables against the number of yolked oocytes as a fixed effect and maternal family as a random effect using the R package nnlme [@Pinheiro2007].

# Results
## Sequencing results
We sequenced a total of 417,846,352 reads across all individuals, 98.8% of which mapped to the monarch reference genome. 244,443,414 reads were maintained after filtering (1,158,500 per sample, on average), encompassing X total sequenced bases. From this, we called `r prettyNum(nsnps(dat),big.mark=",",scientific=FALSE)` individual SNPs following filtering.

@@

Genetic diversity levels were extremely similar between our two years ($\pi =$ `r round(diversity$weighted_mean_pi[2], 3)` and `r diversity$weighted_mean_pi[3]` and $H_{o}$ = `r round(diversity$weighted_mean_ho[2], 3)` and `r round(diversity$weighted_mean_ho[3], 3)` for 2016 and 2018, respectively). Each year had a slight heterozygote excess ($F_{IS}$ = `r round(diversity$weighted_mean_fis[2], 3)` and `r round(diversity$weighted_mean_fis[3],3)` for 2016 and 2018, respectively), although that excess was lower if all samples were considered jointly ($F_{IS}$ = `r round(diversity$weighted_mean_fis[1], 3)`). Pairwise $F_{ST}$ between the two years was small but marginally significant (`r round(fst$weighted_mean_fst[1], 3)`, $p =$ `r round(fst$weighted_mean_fst_p[1], 3)`. We found no significant differences between the two sample years on either the PCA or the NGSadmix results, with clustering instead driven by family (Figure 1).

@@

Diapuse status was strongly biased by Maternal Family (Fisher's Exact Test, $p$ = `r round(fisher_res$p.value)`, see Figure 2a). We identified one small region of chromosome 11 roughly between positions 5,589,278 to 5,619,459 containing many loci associated with diapause status well beyond the background level of association, including three SNPs from two different RAD-tags that had a $p$-value  $\le 5 \times 10^{-5}$ (Figure 2b). This region overlaps the gene DPOGS208560, a protein modifying gene, and was still strongly associated with diapause if only the 2018 butterflies alone were examined. We also identified one locus with a $p$-value  $\le 5 \times 10^{-5}$ on chromosome 16, but there were no other significantly associated loci nearby.

We found that wing shape was significantly negatively correlated with the number of yolked oocytes in Freedman et al.'s [-@Freedman2018] earlier study ($p$ = `r round(summary(mod_AsRa)$tTable[2,5], 3)`) after accounting for maternal family. Wing size was also negatively correlated with the number of yolked oocytes, but not significantly so after accounting for maternal family ($p$ = `r round(summary(mod_Area)$tTable[2,5], 3)`.

@@

We found that a region on Chromosome 11 containing the gene DPOGS208560 was strongly associated with diapuase initiation in Monarch Butterflies in QLD, Australia. Interestingly, a predicted homolog of this gene in silkworms (*Bombyx mori*) has been previously differential expressed during larval molt-intermolt transitions [@Hu2016], which are generally controlled by fluctuations in juvenile hormone (JH) and ecdysone [@Jindra2013]. JH is also known to be a major driver of reproductive development and diapause in monarchs [@Herman2001], so it is not particularly surprising that DPOGS208560 could contribute to diapause control in monarchs.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(snpR); library(data.table); library(ggplot2); library(gridExtra); library(dplyr); library(openxlsx); library(nlme)

@@

Migratory species are acutely at risk from global anthropogenic environmental change, since they rely on multiple distinct ecosystems and consistent seasonal timings [@Both2010]. As a result, the loss of breeding, wintering, or any transitional habitats may all cause declines in migrants [@Robbins1989], as may shifts in resource phenology due to seasonal changes which can cause resource availability and species life-histories to fall out of sync [@Jones2010a]. It is therefore not surprising that widespread declines have been observed in many migratory taxa [@Limburg2009;@Robbins1989].

Population-level adaptability in migratory life-history should therefore be expected to increase long-term population viability by conferring resistance to phenological and environmental changes [@Pulido7341]. Migratory life-history adaptation can essentially occur along three distinct axis: 1) *temporal*, or adaptation that changes the phenology of migration; 2) *spatial*, or adaptation that alters migratory pathways and orientation; and 3) *residency*, or adaptation that shifts populations between migratory and resident life histories. Inter-population variation and adaptation in each of these pathways is well known from nature: temporal changes or variation in migration are well documented in many taxa, particularly in birds [@Rubolini2007; @Gordo2007];  as is short-stopping, extension of migratory pathways, or complete re-direction of migration [@Ambrosini2011; @LaSorte2007; @Visser2009]; and shifts towards residency or partial migration have been observed in birds [@Pulido7341; @Plummer2015; @Møller2014], butterflies [@Howard2010], mammals [@VanDerRee2006; @Fleming2003; @Jones2014a], and other taxa.

Adaptation via a transition to residency is of both particular interest and concern for the conservation of migratory species. While switching to a resident life history can allow species to avoid the increasing risks of migration and the intrinsically very large energetic costs of migration in general [@McWilliams2004; @Rankin1992; @Wikelski2003], there is reasonable concern that the "extinction" of migration may not be easily reversible in some species. This is concerning since migratory life-history variation is expected to increase population adaptability [@Pulido2007], and thus the permanent loss of migration could potentially compromise long-term population persistence. Beyond this, migratory species often provide critical ecosystem services [@lopez-hoffmanEcosystemServicesTransborder2017;@semmensAccountingEcosystemServices2011] and have cultural and/or ecological  significance [@burgerComplexityBioindicatorSelection2015;@schuetzCharacterizingCulturalNiches2019;@semmensQuantifyingEcosystemService2018;@reynoldsMigratorySpeciesEcological2011] that may not survive the transition to permanent residency.

The loss of migratory behavior is a major concern for Monarch Butterflies (*Danaus plexippus*) [@browerAnimalMigrationsEndangered1991;@satterfieldMigratoryMonarchsThat2018]. Specifically, North American migratory monarch populations have experienced severe declines over the last several decades, sparking a fear that those populations may go extinct in the near future [@Semmens2016]. However, many individual butterflies ostensibly from the North American migratory population are now year-round residents in some areas in the southern United States [@howardCitizenScienceObservations2010], where they can survive due to the presence year-round of introduced tropical milkweed (*Asclepias curassavica*) [@bataldenPotentialChangesEastern2015], and there are wide-spread, non-migratory monarch populations that are not of any immediate conservation concern [@walkerDanausPlexippusErrata]. It is specifically, therefore, migration that is at risk in monarch butterflies, not the species as a whole. However, in addition to the damage to the long-term population persistence that could result from the loss of migration in North American populations (as discussed above), the annual monarch migration is publicly beloved and its loss would be culturally costly [@Gustafsson2015].

Interestingly, however, there a population of monarchs in Australia in which migratory behavior has apparently re-emerged after a it was previously lost. This population, which was established sometime before 1871 [@Zalucki2004], is the product of a series of sequential introductions that crossed the Pacific Ocean from North America beginning sometime in the early 1840s after the introduction of their host plants enabled individuals to survive after being blown from the mainland [@Zalucki2004; @Pierce2014; @Hemstrom2022a]. Monarchs are not known to be migratory on any Pacific islands, and no migratory behavior was observed in Australia until the 1930s, when non-reproductive overwintering aggregations were observed in northern New South Wales [NSW, see @James2019]. These and other individuals were later shown to migrate hundreds of kilometers each year to their wintering grounds from more southerly, inland locations in NSW [@James2019;@jamesInductionReproductiveDormancy1983]. After over 90 years (and hundreds of generations) without any apparent migration, the behavior therefore re-emerged in these populations. Understanding how this happened could provide crucial context around the "extinction" of migration in monarchs and other species.

To better understand how migration re-emerged in Australian monarchs, we studied continuously breeding monarch populations in Queensland. While these populations do not experience substantial seasonal fluctuations in resource availability, and thus do not migrate [@James2019], recent work has shown that reproductive diapause can be induced in some, *but not all* individuals from these populations if they are exposed to a reduction in photoperiod during development [@Freedman2018], a trait that these non-migratory populations do not display in the wild. In brief, reproductive diapause, where individuals delay their reproductive maturation, is a critical and tightly linked phenotype for migratory monarchs because it allows them to invest more heavily in lipid storage and increase their average longevity [@Herman2001]. Like other migratory behaviors, diapause is initiated by changes in day length, temperature, and other seasonal fluctuations [@Goehring2002;@iiamsPhotoperiodicClockRegulation2019].

The variance in diapause induction in Queensland monarchs allows for a direct search for genetic associations with this trait and thus for an examination of the genetic underpinnings of the re-evolution of migration in Australian monarchs. *Here, we used reduced representation sequencing to search for genes associated with migratory diapause in Queensland monarch butterflies. We found one strongly associated locus which contained with an E3 ubiquitin ligase gene. This gene has been previously identified as associated with intermolt transitions in silk moths (Bombyx mori), which is controlled by the same major hormone (juvenile hormone, hereafter JH) as reproductive diapause.* This gene had not previously been identified as associated with migration in North American monarch butterflies, and we did not find that any previously identified migratory genes were associated with diapause in our study.

# Methods
## Background
* life history details
* why was migration lost on the islands
* have the map figure here
* why does diapause matter

## Sample Collection, Incubation, and Phenotyping
```{r, read_sample_meta_get_numbers, echo=FALSE,message=FALSE,include=FALSE,warning=FALSE}
sample.meta <- read.table("../data/sample_metadata.txt", header = T, stringsAsFactors = F)
old.data <- data.table::fread("../data/Aus_2018_metadata.txt", header = T)
old.data[,fam := as.numeric(gsub("_.+", "", ID))]

@@

We collected 22 female monarch butterflies from Pinjarra Hills, Queensland, Australia (27°32′26.7″S, 152°54′22.7″E) between the 5th and 9th of July 2018. Individuals in all life stages were present at the field site during our sample collections, consistent with continuous, year-round breeding. We enclosed each of these 22 females individually on either tropical or *Gomphocarpus* species milkweeds found on-location and subsequently successfully collected eggs produced by 21 of them. While these eggs represent 21 total maternal families, each may consist of multiple groups of half-siblings since females may mate multiple times and lay clutches with mixed parentage. We did not remove *Ophryocystis elektroscirrha* (a protozoan parasite) spores, which were uniformly present on our eggs [consistent with previous reports from this location, see @Freedman2018].

We then incubated all eggs using the same incubators and following the "decreasing photoperiod" experimental scheme of Freedman et al. [-@Freedman2018] in order to induce reproductive diapause and ensure that our data could be later pooled with that produced in the earlier study. Briefly, larvae were reared at a constant temperature of 28°C under a photoperiod regime that declined from 14:10 L:D to 12:12 L:D over the course of 30 days (Δ4 mins/day). We released all male butterflies following emergence, since phenotyping diapause in male monarchs can be challenging due to the difficulty in extracting and weighing the ejaculatory duct, seminal vesicle, and accessory glands, which are likely the best indicator of male reproductive development [@Goehring2002]. This left us with a total of `r nrow(old.data)` females from `r length(unique(na.omit(old.data$fam)))` maternal families.

We assessed reproductive diapause by determining the degree of oocyte development in each of our adult female butterflies. We observed that individuals fell into two clearly defined bins: `r sum(old.data$Phenotype == "C", na.rm = TRUE)` out of `r nrow(old.data)` had fully developed, chorionated eggs (visible ridges along the exterior of the egg), while `r sum(old.data$Phenotype != "C", na.rm = TRUE)` females did not. Of these latter individuals, a few (`r sum(old.data$Phenotype == "I", na.rm = TRUE) + sum(old.data$Phenotype == "N", na.rm = TRUE)`) had no yolk in their eggs whatsoever, but most had some degree of yolking. Since vitellogenesis generally only begins to a substantial degree after eclosion in monarchs [@Pan1976], we called any individuals with more than a small amount of yolking in their eggs as reproductively mature (N = `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "M", na.rm = TRUE)` and `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "R", na.rm = TRUE)` for reproductively immature and mature individuals, respectively, images of egg development in each individual are available from the authors upon request). Wings and bodies from all individuals were preserved in dry coin envelops or 95% ethanol, respectively. We note that while we refer to the postponement of reproductive development under other suitable conditions as diapause, other authors refer to this process in monarchs as oligopause [@jamesOVARIANDORMANCYDANAUS1982] or quiescence, since diapause implies an extended refractory period.

We supplemented these samples with preserved, dry butterfly wings from each of the "decreasing photoperiod" female individuals previously scored for reproductive development by Freedman et al. [-@Freedman2018], constituting an additional `r sum(sample.meta$year == 2016, na.rm = TRUE)` total female butterflies from `r length(unique(meta.micah$Female_Line[meta.micah$Treatment == "Decreasing"]))` maternal families, of which `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "M", na.rm = TRUE)` and `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "R", na.rm = TRUE)` were scored by the authors as reproductively immature and mature, respectively. Together, these two datasets contained `r length(na.omit(sample.meta$likely_pheno))` monarch butterflies, of which `r sum(sample.meta$likely_pheno == "M", na.rm = TRUE)` and `r sum(sample.meta$likely_pheno == "R", na.rm = TRUE)` were reproductively immature and mature, respectively (see Table 1).

## Sequencing and Genotyping
We removed and extracted DNA from one leg from each of our samples collected in 2018 or from a wing base from each of the samples collected by Freedman et al [-@Freedman2018] using the magnetic bead protocol of Ali et al [-@Ali389]. We quantified the resulting DNA on a BioTek Instruments FLx800 Fluorescence Reader using Thermo Fisher Scientific Quant-iT PicoGreen dsDNA Reagent, the prepared Restriction Associated Digest (RAD) libraries using the Pst1 restriction enzyme according to Ali et al [-@Ali389]. We sequenced these 150 bp paired-end sequencing libraries using an Illumina Hi-Seq 4000.

We aligned the resulting raw sequence data to the v4 monarch butterfly genome assembly [@Gu2019] using the mem algorithm of the Burrows-Wheeler Aligner [@Li2009a]. We then filtered out PCR duplicates, improperly paired, or poorly mapped reads using SAMtools [@Li2009]. We then called gentoypes using the ANGSD software package with the following parameters: -doMajorMinor 1 (determine major and minor alleles using a genotype likelihood approach),  -doMaf 2 (determine minor allele frequencies), -SNP_pval 1e-8 (keep only loci with a SNP p-value $\le 1 \times 10^{-8}$), -doGeno 4 (call genotypes), -doPost 2 (calculate genotype posterior probabilities using a uniform prior), -postCutoff 0.95 (keep only loci where the highest genotype posterior probability $\ge 0.95$), -minQ 20 (keep only loci with a sequencing quality $\ge 20$), -minMapQ 20 (keep only loci with a mapping quality $\ge 20$), -minInd 130 (keep only loci sequenced in at least 130 individuals), and -minMaf 0.05 (keep only loci with a minor allele frequency $\ge 0.05$).

@@

@@

```{r import_and_diversity, echo=FALSE, message=FALSE, warning=FALSE,include=FALSE}
L <- 6578062 # total number of sequenced bases

@@

#==========diversity===========
# dat <- calc_pi(dat, c(".base", "year"))
# dat <- calc_ho(dat, c(".base", "year"))
# dat <- calc_fis(dat, c(".base", "year"))
# dat <- calc_pairwise_fst(dat, "year", boot = 100, boot_par = 6, verbose = TRUE)
# saveRDS(dat, "diversity_stats.RDS")
dat <- readRDS("diversity_stats.RDS")
diversity <- get.snpR.stats(dat, c(".base", "year"), c("pi", "ho", "fis"))$weighted.means
diversity$weighted_mean_pi <- diversity$weighted_mean_pi/L
diversity$weighted_mean_ho <- diversity$weighted_mean_ho/L
fst <- get.snpR.stats(dat, "year", "fst")$weighted.means

@@

In order to determine if any genomic regions were significantly associated with reproductive diapause, we used the snpR R package to filter out loci significantly out of Hardy-Weinburg Equilibrium ($p \le 0.000001$) in either year in order to help remove any paralogous loci and interpolated missing data by conducting binomial draws at missing genotypes using the minor allele frequencies at those loci [@hemstromSnpRUserFriendly2022]. To assess the basic level of genetic diversity amongst our samples, we then calculated the average number of pairwise differences ($\pi$), observed heterozygosity ($H_{O}$), and $F_{IS}$ according to Weir and Cockerham [-@Weir1984] for all of our samples and for each year independently using snpR [@hemstromSnpRUserFriendly2022]. We also calculated $F_{ST}$ according to Weir and Cockerham [-@Weir1984] between sample years and between both sample year and reproductive phenotype jointly [@hemstromSnpRUserFriendly2022]. We bootstrapped a significance level for $F_{ST}$ by bootstrapping individuals among populations with replacement 100 times. To determine if there was any substantial unexpected genetic structuring (other than family groups) in our samples, we conducted a Principal Component Analysis using snpR [@hemstromSnpRUserFriendly2022]. Since some individuals escaped their rearing dishes (but not the incubator) during rearing, we also confirmed the parentage of our samples using the Colony2 program, assuming polygamous, random mating [@Jones2010].

@@

# plot
Figure2b <- plot_manhattan(dat, "gmmat_pval_likely_pheno", chr = "scaffold", significant = 0.00005, suggestive = 0.0005, log.p = T,
                           chr_order = c(paste0("chr", 1:30), "other"), highlight_style = "red") # highlight top snp
Figure2b <- Figure2b$plot + ylab("-log10(p-value)") + xlab("Position") + theme(strip.text = element_blank())
```
In order to determine the degree to which diapause status was biased between families, we constructed a Fisher's exact test for independence using 2000 simulations in R version 4.2.2 [@rcoreteamLanguageEnvironmentStatistical2022]. We then conducted a Genome-Wide Association Study (GWAS) in order to determine if any loci were significantly associated with reproductive status. Since our samples were composed of many groups of full and half-siblings and came from two genetically different years, we used the GMMAT R package to conduct a population and family structure-corrected GWAS [@Chen2020; @Chen2016]. We visualized the results by constructing a Manhattan plot using snpR [@hemstromSnpRUserFriendly2022]. We identified the genes underlying any candidate regions by hand using the MonarchBase genome browser online tool [@Zhan2013].

@@

In order to determine if diapause is part of a generalized migratory syndrome in Australia, we also briefly re-analyzed the connection between reproductive status and wing morphology in our 2016 samples [@Freedman2018]. In order to determine if wing shape or size was correlated with reproductive status, we fit a pair of linear mixed effect models with either with wing shape or size as response variables, the number of yolked oocytes as a fixed effect, and maternal family as a random effect using the R package nnlme [@Pinheiro2007].

# Results
## Sequencing results
We sequenced a total of 417,846,352 reads across all individuals, 98.8% of which mapped to the monarch reference genome. 244,443,414 reads were maintained after filtering (1,158,500 per sample, on average), encompassing `r prettyNum(L,big.mark=",",scientific=FALSE)` total sequenced bases. From this, we called `r prettyNum(nsnps(dat),big.mark=",",scientific=FALSE)` individual SNPs following filtering.

@@

Genetic diversity levels were extremely similar between our two years ($\pi =$ `r round(diversity$weighted_mean_pi[2], 4)` and `r round(diversity$weighted_mean_pi[3], 4)` and $H_{o}$ = `r round(diversity$weighted_mean_ho[2], 4)` and `r round(diversity$weighted_mean_ho[3], 4)` for 2016 and 2018, respectively). Each year had a slight heterozygote excess ($F_{IS}$ = `r round(diversity$weighted_mean_fis[2], 3)` and `r round(diversity$weighted_mean_fis[3],3)` for 2016 and 2018, respectively), although that excess was lower if all samples were considered jointly ($F_{IS}$ = `r round(diversity$weighted_mean_fis[1], 3)`). Pairwise $F_{ST}$ between the two years was small but marginally significant (`r round(fst$weighted_mean_fst[1], 3)`, $p =$ `r round(fst$weighted_mean_fst_p[1], 3)`). We found no significant genetic divergence between the two sample years on either the PCA or the NGSadmix results, with clustering instead driven by family, although we did find that three of our 2018 maternal families were substantially different from all other samples on our first two PC axes (Figure 1).

@@

Diapause status was strongly biased by maternal family (Fisher's Exact Test, $p < 0.001 $, see Figure 2a). We identified one small region of chromosome 11 (between positions ~5,589,000 to ~5,620,000) containing many loci associated with diapause status well beyond the background level of association, including three SNPs from two different RAD-tags that had a $p$-value  $\le 5 \times 10^{-5}$ (Figure 2b). This region overlaps the gene DPOGS208560 (LOC116768596), an E3 ubiquitin ligase, and was still strongly associated with diapause if only the 2018 butterflies alone were examined. We also identified one locus with a $p$-value  $\le 5 \times 10^{-5}$ on chromosome 16, but there were no other significantly associated loci nearby.

We found that wing shape was significantly negatively correlated with the number of yolked oocytes in Freedman et al.'s [-@Freedman2018] earlier study ($p$ = `r round(summary(mod_AsRa)$tTable[2,5], 3)`) after accounting for maternal family. Wing size was also negatively correlated with the number of yolked oocytes, but not significantly so after accounting for maternal family ($p$ = `r round(summary(mod_Area)$tTable[2,5], 3)`).

@@

We found that a region on Chromosome 11 containing the gene DPOGS208560 was strongly associated with diapause initiation in Monarch Butterflies in Queensland, Australia. This gene is an E3 ubiquitin ligase, a class of proteins broadly involved with proteosomal degradation of diverse substrates, including the circadian clock protein CRY2 [@xingSCFFBXL3UbiquitinLigase2013; @yooCompetingE3Ubiquitin2013], which is known to play a role in monarch photoperiodism including migratory responses to declining photoperiod [@iiamsPhotoperiodicClockRegulation2019]. Interestingly, a predicted homolog of this gene in silk moths has been previously found to be differentially expressed during larval molt-intermolt transitions [@Hu2016], which are generally controlled by fluctuations in JH and ecdysone [@Jindra2013]. JH is also known to be a major driver of reproductive development and diapause in monarchs [@Herman2001]. Taken together, it is not surprising that DPOGS208560 may participate in diapause control in monarchs.
